---
title: "Part A – SQL Queries"
format: html
execute:
  echo: true
---

# Part A – SQL Queries

This section demonstrates the use of SQL to query and analyse an SQLite database containing information about interactive cinematic virtual production projects.  
The database includes productions, clients, environments, assets, crew members, shooting sessions, and equipment.

All queries are executed using R with the `DBI` and `RSQLite` packages, and results are displayed using interactive tables.

## Database Connection and Exploration

We begin by connecting to the SQLite database and inspecting its structure.  
Listing the available tables and key fields ensures that queries reference the correct table and column names.


```{r}
library(DBI)
library(RSQLite)
library(DT)

con <- dbConnect(
  SQLite(),
  "data/virtual_production.db"
)
```

The database contains tables related to productions, clients, crew members, assets, environments, shooting sessions, and equipment.  
This confirms that the dataset supports analysis across financial, temporal, and personnel dimensions of virtual production.

## Query 1 – Total Budget by Country

**Objective:**  
List the total budget allocated for productions in each country, along with the number of projects per country.  
The results are sorted by total budget in descending order.

**Approach:**  
- Join the `Clients` and `Productions` tables using `ClientID`
- Aggregate production counts and budgets by country
- Use a `LEFT JOIN` to ensure countries with no productions are included


```{r}
query1 <- "
SELECT 
    c.Country,
    COUNT(p.ProductionID) AS ProjectCount,
    COALESCE(SUM(p.Budget), 0) AS TotalBudget
FROM Clients c
LEFT JOIN Productions p
    ON c.ClientID = p.ClientID
GROUP BY c.Country
ORDER BY TotalBudget DESC;
"

result1 <- dbGetQuery(con, query1)
DT::datatable(result1, options = list(pageLength = 8))



```

**Result Interpretation:**

- Germany, the UK, and Japan have the highest total production budgets.
- Some countries (e.g. USA and Canada) appear in the client list but currently have no associated productions.
- This query demonstrates aggregation (`COUNT`, `SUM`), joins, grouping, and sorting.


## Query 2 – Average Development Time by Asset Count

**Objective:**  
Calculate the average development time for productions, grouped by the number of assets used.

**Approach:**  
- Count the number of assets per production via `Environments` and `Assets`
- Calculate development time using the difference between `StartDate` and `EndDate`
- Group results by asset count to identify trends

```{r}
query2 <- "
SELECT 
    AssetCounts.AssetTotal AS NumberOfAssets,
    AVG(julianday(p.EndDate) - julianday(p.StartDate)) AS AvgDevelopmentDays
FROM Productions p
JOIN (
    SELECT 
        e.ProductionID,
        COUNT(a.AssetID) AS AssetTotal
    FROM Environments e
    JOIN Assets a
        ON e.EnvironmentID = a.EnvironmentID
    GROUP BY e.ProductionID
) AS AssetCounts
    ON p.ProductionID = AssetCounts.ProductionID
GROUP BY AssetCounts.AssetTotal
ORDER BY NumberOfAssets;
"

result2 <- dbGetQuery(con, query2)
DT::datatable(result2, options = list(pageLength = 8))


```

**Result Interpretation:**

- Productions with a higher number of assets tend to show longer development times.
- However, the relationship is not strictly linear, suggesting other factors such as production complexity or team efficiency may influence timelines.
- This query demonstrates subqueries, joins, date calculations, and aggregation.


## Query 3 – Top Crew Members by Successful Productions

**Objective:**  
Identify the top three crew members based on the number of successful productions they have participated in.

**Definition of Success:**  
Productions with a status of **Post-production** or **Completed** are considered successful.

**Approach:**  
- Join `CrewMembers`, `ProductionCrew`, and `Productions`
- Filter productions by status
- Count the number of successful productions per crew member
- Sort results and limit output to the top three


```{r}
query3 <- "
SELECT 
    cm.Name,
    cm.Role,
    COUNT(pc.ProductionID) AS SuccessfulProductions
FROM CrewMembers cm
JOIN ProductionCrew pc
    ON cm.CrewID = pc.CrewID
JOIN Productions p
    ON pc.ProductionID = p.ProductionID
WHERE p.Status IN ('Post-production', 'Completed')
GROUP BY cm.CrewID, cm.Name, cm.Role
ORDER BY SuccessfulProductions DESC
LIMIT 3;
"

result3 <- dbGetQuery(con, query3)
DT::datatable(result3, options = list(pageLength = 8))


```

**Result Interpretation:**

- Unreal Operators appear prominently among the most successful crew members.
- This highlights the importance of technical Unreal Engine roles in virtual production pipelines.
- This query demonstrates filtering, grouping, ordering, and limiting results.


## Demonstration of Core SQL Concepts

The following queries demonstrate key SQL concepts as required by the assignment.

### SELECT with LIKE and OR

**Objective:**  
Retrieve crew members whose roles include either *Operator* or *Specialist*.

**Approach:**  
- Use the `LIKE` operator with wildcards
- Combine multiple conditions using `OR`


```{r}
query4 <- "
SELECT Name, Role
FROM CrewMembers
WHERE Role LIKE '%Operator%'
   OR Role LIKE '%Specialist%';
"

result4 <- dbGetQuery(con, query4)
DT::datatable(result4, options = list(pageLength = 8))

```

This query demonstrates pattern matching and logical conditions in SQL.

### SELECT with DISTINCT and ORDER BY

**Objective:**  
List all unique Unreal Engine versions used across environments, sorted in descending order.

**Approach:**  
- Use `DISTINCT` to remove duplicates
- Sort results using `ORDER BY`


```{r}
query5 <- "
SELECT DISTINCT EngineVersion
FROM Environments
ORDER BY EngineVersion DESC;
"

result5 <- dbGetQuery(con, query5)
DT::datatable(result5, options = list(pageLength = 8))


```

This query highlights version diversity and demonstrates de-duplication and sorting.

### Subquery with SELECT

**Objective:**  
Identify productions with a budget higher than the average production budget.

**Approach:**  
- Calculate the average budget using a subquery
- Compare each production’s budget against this value

```{r}
query6 <- "
SELECT Title, Budget
FROM Productions
WHERE Budget > (
    SELECT AVG(Budget)
    FROM Productions
);
"

result6 <- dbGetQuery(con, query6)
DT::datatable(result6, options = list(pageLength = 8))

```

This query demonstrates the use of subqueries for comparative analysis.

## Conclusion

Part A demonstrates the effective use of SQL for data exploration and analysis within a virtual production context.  
The queries showcase joins, aggregations, filtering, subqueries, and sorting, providing meaningful insights into budgets, timelines, assets, and crew involvement.

The use of interactive tables enhances readability and allows users to explore the results dynamically.



```{r}
dbDisconnect(con)

```
