---
title: "Part B – Linear Regression Analysis"
format: html
execute:
  echo: true
---

# Linear Regression Analysis of Cinematic Productions

In this section, we perform a linear regression analysis to understand how production **budget** and **crew size** affect the **success rate** of a production.

The **success rate** is defined as:

\[
\text{SuccessRate} = \frac{\text{ExpectedDuration}}{\text{ActualWorkload}}
\]

Where:

- **ExpectedDuration** is the planned duration of the production (`EndDate - StartDate`)  
- **ActualWorkload** is the total duration of all shooting sessions.

## Data Preparation

We first calculate the total actual workload for each production by summing the duration of all shooting sessions.  
We then merge this with the production budget and crew size to create the dataset for modeling.

```{r}
library(DBI)
library(RSQLite)
library(tidyverse)
library(ggplot2)

# Connect to database
con <- dbConnect(SQLite(), "data/virtual_production.db")

# Sum duration hours per production
actual_workload <- dbGetQuery(con, "
SELECT 
    ProductionID,
    SUM(DurationHours) AS ActualWorkload
FROM ShootingSessions  -- correct plural table
GROUP BY ProductionID
")

# Get production info and crew size
production_info <- dbGetQuery(con, "
SELECT 
    p.ProductionID,
    p.Budget,
    julianday(p.EndDate) - julianday(p.StartDate) AS ExpectedDuration,
    COUNT(pc.CrewID) AS CrewSize
FROM Productions p
LEFT JOIN ProductionCrew pc
    ON p.ProductionID = pc.ProductionID
GROUP BY p.ProductionID
")

# Merge datasets and calculate success rate
data_model <- production_info %>%
  left_join(actual_workload, by = "ProductionID") %>%
  mutate(
    SuccessRate = ExpectedDuration / ActualWorkload
  )

head(data_model)

```

## Linear Regression Model

We fit a linear regression model to predict `SuccessRate` based on **Budget** and **CrewSize**.

The model coefficients will help us understand the relationship between these variables and production success.

```{r}
# Fit linear regression
model <- lm(SuccessRate ~ Budget + CrewSize, data = data_model)

# Show model summary
summary(model)

```

## Model Visualization

We plot the success rate against budget and crew size to visualize trends and patterns.

```{r}
ggplot(data_model, aes(x = Budget, y = SuccessRate, color = CrewSize)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Success Rate vs Budget & Crew Size",
    x = "Budget",
    y = "Success Rate",
    color = "Crew Size"
  ) +
  theme_minimal()

```

## Model Interpretation

- **Intercept**: the predicted success rate when both budget and crew size are zero.
- **Budget coefficient**: the change in success rate for each unit increase in budget, keeping crew size constant.
- **CrewSize coefficient**: the change in success rate for each additional crew member, keeping budget constant.

**Insights**:

- A positive coefficient indicates that increasing the variable improves the success rate.
- A negative coefficient indicates that increasing the variable reduces the success rate.

## Model Reliability & Outliers

We examine the residuals and leverage points to identify outliers and evaluate the reliability of the model.

```{r}
# Check residuals and model diagnostics
par(mfrow = c(2,2))
plot(model)

```

Observations:

- Residual vs Fitted: should be randomly scattered; patterns may indicate non-linearity.
- Normal Q-Q: checks residual normality.
- Scale-Location: checks for homoscedasticity.
- Residuals vs Leverage: identifies influential outliers.

Conclusion: Linear regression provides a good first approximation but may be sensitive to outliers or extreme values in budget or crew size.

```{r}
dbDisconnect(con)

```
